My title

### 2024-04-19 paper/paper.qmd
Discussion section

---
title: "My title"
author: 
  - Carl Fernandes
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "This study investigates the connection between income inequality and national hapiness, questioning the impact of the Gini index. By examining cross-sectional data we find that income inequality has a limited direct effect on happiness. It is evident that factors like the quality of institutions and levels of development as measured by the Democracy Index and the Inequality adjusted Human Development Index (IHDI) have a more significant impact on shaping happiness. Our results indicate that focusing solely on income inequality doesn't provide a complete understanding. Instead a wider range of indicators are. This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

## Background
Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress. This paper embarks on an empirical inquiry into the determinants of happiness, with a particular focus on income distribution. Traditional economic theories suggest a correlation between economic prosperity and happiness, but emerging research prompts a re-evaluation of this association. In this context, income inequality, typically measured by the Gini index, emerges as a potential factor influencing happiness. Yet, the depth of its impact remains under-examined and is often overshadowed by more visible economic indicators.

@@

### Simple Linear Model
The initial model considers the Gini index as the sole predictor. This setup allows for an assessment of how variations in income inequality might align with changes in happiness scores. It is represented by the equation below where $\alpha$ is the intercept, $\beta$ is the slope coefficient for the Gini index, and $\epsilon_i$ is the error term.

```{=tex}
$$
\begin{aligned}
HappinessScore_{i} &= \alpha + \beta \times GiniIndex_{i} + \epsilon_{i} \\
\epsilon_{i} &\sim N(0, \sigma^2)
\end{aligned}
$$
```

### Multi-Linear Model

@@

```{=tex}
\begin{align} 
\text{Happiness\_score}_i &= \alpha + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
\varepsilon &\sim \mbox{Normal}(0, \sigma^2)
\end{align}
```

Both models are estimated using ordinary least squares (OLS), which minimizes the sum of squared differences between the observed and predicted values. This method was chosen for its robustness and simplicity, facilitating the interpretation of coefficient estimates.

@@

The extended model combines the Democracy index and IHDI with the Gini index to explore how they collectively influence happiness levels. The findings reveal a connection between IHDI and happiness scores ($\beta_3$ = 4.112, p < 0.001) indicating that nations with higher human development levels, adjusted for inequality show higher happiness among their populace.

@@

## Diagnostic checks
The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution. This deviation from the expected straight line in the Q-Q plot could imply that the residuals do not follow a perfect normal distribution, which is an assumption for the OLS regression. However, the majority of the points do align well with the theoretical quantiles, indicating that any deviations from normality are not extreme.

@@

# Model details {#sec-model-details}
In preparation for the OLS regression, the dataset was meticulously inspected for multicollinearity to prevent inflated variance in the coefficient estimates. The IHDI, adjusted for income disparities, was chosen specifically to avoid overlap with the Gini index. The assumptions of residual normality, constant variance across predictions (homoscedasticity), and the independence of observations were each tested. Any deviations from these conditions prompted corrective measures such as variable transformations to achieve normality and homoscedasticity, and the use of robust standard errors to counteract potential heteroscedasticity, ensuring the robustness of our regression results.

>>>>

---
title: "My title"
author: 
  - Carl Fernandes
thanks: "Code and data are available at: https://github.com/Carl-1506/Happiness-and-inequality.git"
date: today
date-format: long
abstract: "This study investigates the connection between income inequality and national hapiness, questioning the impact of the Gini index. By examining cross-sectional data we find that income inequality has a limited direct effect on happiness. It is evident that factors like the quality of institutions and levels of development as measured by the Democracy Index and the Inequality adjusted Human Development Index (IHDI) have a more significant impact on shaping happiness. Our results indicate that focusing solely on income inequality doesn't provide a complete understanding. Instead a wider range of indicators are. This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

## Background

Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress. This paper embarks on an empirical inquiry into the determinants of happiness, with a particular focus on income distribution. Traditional economic theories suggest a correlation between economic prosperity and happiness, but emerging research prompts a re-evaluation of this association. In this context, income inequality, typically measured by the Gini index, emerges as a potential factor influencing happiness. Yet, the depth of its impact remains under-examined and is often overshadowed by more visible economic indicators.

@@

### Simple Linear Model

The initial model considers the Gini index as the sole predictor. This setup allows for an assessment of how variations in income inequality might align with changes in happiness scores. It is represented by the equation below where $\alpha$ is the intercept, $\beta$ is the slope coefficient for the Gini index, and $\epsilon_i$ is the error term.

```{=tex}
$$
\begin{aligned}
HappinessScore_{i} &= \alpha + \beta \times GiniIndex_{i} + \epsilon_{i} \\
\epsilon_{i} &\sim N(0, \sigma^2)
\end{aligned}
$$
```
### Multi-Linear Model

@@

```{=tex}
\begin{align} 
\text{Happiness\_score}_i &= \alpha + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
\varepsilon &\sim \mbox{Normal}(0, \sigma^2)
\end{align}
```
Both models are estimated using ordinary least squares (OLS), which minimizes the sum of squared differences between the observed and predicted values. This method was chosen for its robustness and simplicity, facilitating the interpretation of coefficient estimates.

@@

The extended model combines the Democracy index and IHDI with the Gini index to explore how they collectively influence happiness levels. The findings reveal a connection between IHDI and happiness scores ($\beta_3$ = 4.112, p \< 0.001) indicating that nations with higher human development levels, adjusted for inequality show higher happiness among their populace.

@@

## Diagnostic checks

The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution. This deviation from the expected straight line in the Q-Q plot could imply that the residuals do not follow a perfect normal distribution, which is an assumption for the OLS regression. However, the majority of the points do align well with the theoretical quantiles, indicating that any deviations from normality are not extreme.

@@

# Model details {#sec-model-details}

In preparation for the OLS regression, the dataset was meticulously inspected for multicollinearity to prevent inflated variance in the coefficient estimates. The IHDI, adjusted for income disparities, was chosen specifically to avoid overlap with the Gini index. The assumptions of residual normality, constant variance across predictions (homoscedasticity), and the independence of observations were each tested. Any deviations from these conditions prompted corrective measures such as variable transformations to achieve normality and homoscedasticity, and the use of robust standard errors to counteract potential heteroscedasticity, ensuring the robustness of our regression results.
